\section{Instructions}
\label{sec:intructions}


\subsection{Getting started}
\label{subsec:getting_started}

\subsubsection{Pulling from GitHub}
\label{subsubsec:pulling_from_github}

The code for this suite can be found at 
\begin{center}
\href{https://github.com/InFoMM/TSP}{https://github.com/InFoMM/TSP}
\end{center}
which should be cloned/forked into whichever directory the user wants to call the functions from.

\subsubsection{Git branches}
\label{subsubsec:git_branches}

To see which branches are available type \\
\indent
\verb|$ git branch| \\
into the console. To switch to one of these branches, type \\
\indent
\verb|$ git checkout <available branch name>|. \\
This is because many of the authors have to present this work, and so ensures submitted work is separate from the \texttt{master} branch.

\subsection{Directory structure}
\label{subsec:directory_structure}

Once the directory has been cloned it should have the following directory structure:
\\
\noindent
\begin{minipage}{\linewidth}
\begin{verbatim}
    TSP/
        matlab/
            increasing_loop.m
            performance.m
            etc ...
        user_guide/
            figures/
            latex/
                guide.pdf
                etc ...
\end{verbatim}
\end{minipage}
\\
\noindent

\subsubsection{Working paths}
\label{subsubsec:working_paths}

To use the functions in \texttt{TSP/matlab} change the Matlab current working directory into this folder. \textbf{Do not add this folder or sub-folders to the working path!} This is because many of these functions use relative path location descriptions to find supporting functions. 

\subsection{Dependencies}
\label{subsec:dependencies}

Unfortunately not all the dependencies are known, but on MATLABR2016b these likely include:
\begin{itemize}
	\item Optimisation Toolbox.
	\item Statistics and Machine Learning.
\end{itemize}
If any others are found please let us know via GitHub.


\subsection{Testing it works}
\label{subsec:testing_it_works}

The directory \texttt{Matlab/} contains the file \verb|test.m| which when run should output the following: \\
\noindent
\begin{verbatim}
Performance of:
        Purmuation searching algorithm.
Time taken:
    2.686500e-01
Total distance:
    2999
Paths taken:
    [1, 1, 10, 3, 2, 5, 6, 9, 8, 7]
Paths taken:
    [4, 4, 7, 8, 9, 6, 5, 2, 3, 10]
Routes taken:
    [Alicante, Alicante, Valencia, Granada, Barcelona, ...
     Malaga, Pamplona, Sevilla, Santander, Salamanca]
Routes taken:
    [Madrid, Madrid, Salamanca, Santander, Sevilla, ...
    Pamplona, Malaga, Barcelona, Granada, Valencia]
\end{verbatim}
where the time taken to complete may be different to the quoted value. 

\subsubsection{Trying various algorithms}
\label{subsubsec:trying_various_algorithms}

To see the performance of some different algorithms see \verb|performance.m|, where we have presented some example code to demonstrate some performance characteristics of various algorithms.

\clearpage